---
title: "Programming Homework 4"
author: "Trever Yoder"
format: pdf
editor: visual
---

### Load Packages

```{r}
library(tidyverse)
```


## Task 1: Conceptual Questions

```{r}
#Create a list with the requested questions
my_list1 <- list("1. What is the purpose of the lapply() function? 
                 What is the equivalent purrr function?",
"2. Suppose we have a list called my_list.
Each element of the list is a numeric data frame (all columns
are numeric). We want use lapply() to run the code
cor(numeric_matrix, method = kendall)
on each element of the list. Write code to do this below!
(I’m really trying to ask you how you specify
method = kendall when calling lapply())",
"3. What are two advantages of using purrr functions
instead of the BaseR apply family?",
"4. What is a side-effect function?",
"5. Why can you name a variable sd in a function and 
not cause any issues with the sd function?")

#print the questions/list
my_list1
```

### Question 1

lapply() is used to apply functions across many rows/columns and the "l" ensures R always outputs a list.

### Question 2

```{r}
#create data frames
df1 <- data.frame(a = c(1:3), b = c(4:6))
df2 <- data.frame(x = c(7:9), y = c(10:12), z = c(13:15))

#create list
my_list <- list(df1, df2)

#write/show lapply function 
lapply(my_list, cor, method = "kendall")
```

### Question 3

Two advantages of using `purr` functions instead of `BaseR` apply family are: 1. Greater consistency between functions. For example, you can predict the output type exclusively from the function name, which isn't always true for `BaseR` apply functions. 2. `Purr` also has some functions to fill in some gaps such as `imap()` where you can map simultaneously over `x` and its indices.

### Question 4

A side-effect function does something beyond it's function return value. For example, write files to a disk. If we want the side effect of `hist` (which is the visual part) we can use the `walk()` function to only print the histogram.

### Question 5

We can name a variable `sd` in a function and not cause any issues with the `sd` function because functions have their own temporary environment. So, once the function executes, all variables within that function are "gone" or in other words, not saved to the main environment.


## Task 2: Writing R Functions

### Question 1

Here I will create a function that calculates RMSE. There is an ellipses in the function to allow for additional arguments.

```{r}
getRMSE <- function(response, predicted, ...){
  add <- list(...)
  remove <- isTRUE(add$na.rm)
  if (remove) {
    dropmissing <- !is.na(response) & !is.na(predicted)
    response <- response[dropmissing]
    predicted <- predicted[dropmissing]
  }
  sqrt(mean((response-predicted)^2))
}
```


### Question 2

Let's run some code to create some response values and predictions.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

#now let's test our RMSE function using this data!
getRMSE(resp, pred)

#Add 2 missing vlaues to resp
resp[c(1,5)] <- NA_real_

#Test with and without specifying what R should do with the missing values
getRMSE(resp, pred)
getRMSE(resp, pred, na.rm = TRUE)
```


### Question 3

Let's create a similar function, except it calculates the MAE instead of the RMSE.

```{r}
getMAE <- function(response, predicted, ...){
  add <- list(...)
  remove <- isTRUE(add$na.rm)
  if (remove) {
    dropmissing <- !is.na(response) & !is.na(predicted)
    response <- response[dropmissing]
    predicted <- predicted[dropmissing]
  }
  mean(abs(response-predicted))
}
```

### Question 4

Now let's create some data and test our MAE function using the data.
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

#Let's test our MAE function
getMAE(resp, pred)

#Add 2 missing vlaues to resp
resp[c(1,5)] <- NA_real_

#Test with and without specifying what R should do with the missing values
getMAE(resp, pred)
getMAE(resp, pred, na.rm = TRUE)
```


### Question 5

Now we want to create a wrapper function that can be used to get either both metrics or a single functioned called.

```{r}
#Create the wrapper function
my_wrapper <- function(response, predicted, metrics = c("MAE", "RMSE"), ...) {
  if (!(is.vector(response) && is.atomic(response) && is.numeric(response))) {
    message("Error: 'response' must be a numeric atomic vecotr.")
    return(invisible(NULL))
  }
  if (!(is.vector(predicted) && is.atomic(predicted) && is.numeric(predicted))) {
    message("Error: 'predicted' must be a numeric atomic vecotr.")
    return(invisible(NULL))
  }
  
  #create empty list for results
  results <- list()
  
  #calculate metrics
  if ("MAE" %in% toupper(metrics)) {
    results$MAE <- getMAE(response, predicted, ...)
  }
  if ("RMSE" %in% toupper(metrics)) {
    results$RMSE <- getRMSE(response, predicted, ...)
  }
  return(results)
}
  
```

### Question 6

Now let's test our wrapper function in a similar way that we tested our previous functions.
```{r}
#create data
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

#call individually then together
my_wrapper(resp, pred, metrics = "MAE")
my_wrapper(resp, pred, metrics = "RMSE")
my_wrapper(resp, pred, metrics = c("MAE", "RMSE"))

#test again but with 2 NA values in resp
resp[c(1,5)] <- NA_real_
my_wrapper(resp, pred, metrics = "MAE")
my_wrapper(resp, pred, metrics = "RMSE", na.rm = TRUE) #tested excluding NA
my_wrapper(resp, pred, metrics = c("MAE", "RMSE"))

#Test by passing a data frame created in Task 1 Question 2.
my_wrapper(df1, df1, metrics = c("MAE", "RMSE"))
my_wrapper(resp, df1, metrics = c("MAE", "RMSE"), na.rm = TRUE)
```

## Task 3: Querying an API and a Tidy-Style Function

### Question 1


### Question 2


### Question 3



```{r}

```
